---
title: "Brilliant Biogeography Wrange Wrangling"
author: "Chauncey Gadek" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 5
    code_folding: show
    bibliography: true
    #df_print: paged
    #df_print: kable
    #toc_float: true
      #collapsed: false
      #smooth_scroll: TRUE
    theme: cosmo #spacelab #yeti #united #cosmo
    highlight: tango
  pdf_document:
    df_print: kable
fontsize: 12pt
geometry: margin=0.25in
always_allow_html: yes
---

```{=html}
<style>
/* HTML FORMATTING */
h1, .h1, h2, .h2, h3, .h3, h4, .h4, h5, .h5 {
  margin-top: 25px; /* space before each header */
  font-weight: bold; /* bold headers */
}
</style>
```
```{R, echo=FALSE}
# I set some GLOBAL R chunk options here.
#   (to hide this message add "echo=FALSE" to the code chunk options)
rm(list =ls (all = TRUE)) #This removes objects from global environ
knitr::opts_chunk$set(echo=F, comment = NA, message = FALSE, warning = FALSE, width = 100)
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 6)

```

# Load Packages

```{R, echo=F}
library(reshape)
library(reshape2)
library(plyr)
library(tidyr)
library(gt)
library(glue)
library(dplyr)
library(scales)
library(letsR)
library(car)
library(rcompanion)
library(GGally)
library(Hmisc)
library(gridExtra)
library(stats)
library(gplots)
library(ggExtra)
library(cowplot)
library(wesanderson)
library(colorspace)
library(stats4) # Forces knitr to work when it's being wonky
library(PMCMR) #Allows Kruskal-Wallis post-hocs
library(effects)
library(gridExtra)
library(lattice)
library(survival)
library(fmsb)
library(faraway)
library(ape)
library(data.table)
#library(wBoot)
library(ggridges)
library(boot)
library(faux)
library(effsize)
library(plotrix)
library(colorspace)
library(ggpubr)
library(patchwork)
library(ggdist)
library(factoextra)
library(tidyverse)
library(readxl)


# Mapping 
library(raster)
library(sp)
library(rgdal)
library(RStoolbox)
library(prettymapr)
library(viridis)
library(rasterVis)
library(maptools)
library(sf)

# Modeling packages 
library(nlme)
library(lme4)
library(AICcmodavg)
library(MuMIn)
library(glmulti)
library(reghelper)
library(lsmeans)
library(rsq) # get r-squared values from GLM
library(r2glmm) # for R^2 values from lmer() and glmer()
library(multcompView) # related to multiple comparisons?
library(jtools) # interaction plots 
library(interactions) # interaction plots 
library(broom)
library(stargazer) # model output tables
library(ggeffects) # for estimating model predictions from mixed effects models
library(MCMCglmm)
library(bayesplot)
library(rstan)
library(Rcpp) # required for brms
library(brms)
library(magrittr)
library(tidybayes)
library(modelr)
library(hexbin)
library(ggExtra)
library(rgl)
library(readr)
library(tidyverse)
# Install remotes if not previously installed
if(!"remotes" %in% installed.packages()[,"Package"]) install.packages("remotes")

# Install rasterSp from Github if not previously installed
if(!"rasterSp" %in% installed.packages()[,"Package"]) remotes::install_github("RS-eco/rasterSp", build_vignettes = T)
library(rasterSp)
# To run each time you load rstan
options(mc.cores = parallel::detectCores()) # for core setup 
rstan_options(auto_write = TRUE) # auto save  bare version of compiled Stan program to HD

#Load in functions
source("~/Dropbox/Gadek_custom_colors.R")
source("~/Dropbox/ggplot_themes/ggplot_themes.R")
#source(paste(getwd(),"/pal.R", sep=""))
data("Phyllomedusa")

#set theme
theme_set(theme_arial_clean())

#setup folder paths for less coding
figures <- paste(getwd(), "/figures/", sep="")
tables <- paste(getwd(), "/Tables/", sep="")
models <- paste(getwd(), "/models/", sep="")
results <- paste(getwd(), "/models/results/", sep="")
```

# Step 1. Download spatial data 
```{r spatial data, message=F}
countries = c("argentina", "bolivia", "brazil", "chile", "colombia", "ecuador", "guyana", "paraguay", "peru", "suriname", "uruguay", "venezuela")
SA = do.call("merge", lapply(countries, function(x) raster::getData('alt', country=x, level=0,path="mapping/")))
projection(SA) <- CRS("+init=epsg:4326") #Make sure raster and spp projections match
plot(SA)

```
 
# Step 2. Species Data 
Full SA BirdLife International dataset 2018 version comes in.dgb file and requires a few extra steps to use. Namely making sure the "All_Species" layer is present.
```{r species data}


#get list of all shape files in folder
file_list <- list.files("mapping/Shapes_True_Brilliants", pattern = "*shp", full.names = TRUE)

#import shapefiles
shapefile_list <- lapply(file_list, read_sf)

# for(i in 1:length(shapefile_list)){
#   names(shapefile_list[[i]]) <-"SCINAME"
# }

#merge them
True_brill <- SpatialPolygons(shapefile_list)

data<-readOGR(dsn=gdb_file, layer= "Distribuciones_True_Brilliants")

tbnames <- read_csv("data/True_brill_names.csv")
tbnames <-tbnames%>%
  pull()



# The input file geodatabase
#gdb_file =here::here("mapping", "Shapes_True_Brilliants")
gdb_file = paste0("~/Dropbox/Research/", "/BOTW.gdb")

# List all feature classes in a file geodatabase
subset(ogrDrivers(), grepl("GDB", name))
fc_list = ogrListLayers(gdb_file)
print(fc_list)

data <- st_read(gdb_file, layer ="All_Species" ) #This takes a long time, huge file!!!!
# Read the feature class
#data <- st_read(gdb_file, layer ="All_Species" ) #This takes a long time, huge file!!!!

colnames(data)

data2 <- data[,c(2,5,6,7)] # We need to keep columns 2,5,6,7 SCINAME, PRESENCE, ORIGIN, SEASONAL in order to use lets.presab
colnames(data2)
head(data2)

# Save as shapefile
#writeOGR(botw, dsn=paste0("~/Dropbox/Research", "/BirdLife_2018/"), layer="All_Species", driver="ESRI Shapefile")
#botw <- spTransform(botw, CRS("+init=epsg:4326"))  #Make sure raster and spp projections match
```


```{r}
#replace spaces in shape_file species names with underscore

tbnames<-gsub('[[:digit:]]+', '', tbnames)
data2$SCINAME  <- sub(" ", "_", data2$SCINAME)

#Load tree
brilliant_tree <- read.nexus("trees/Ster_Hylo_P11.trees")
brilliant_tree <- sample(brilliant_tree, 1)[[1]]

#remove numbers from tree tip labels
brilliant_tree$tip.label <-gsub('[[:digit:]]+', '', brilliant_tree$tip.label)

#some very annoying substrings here (in tree) that need to be replaced manually

brilliant_tree$tip.label<-gsub("rubrfmnh", "rubricauda", brilliant_tree$tip.label) #Clytolaema_rubrnmh"
brilliant_tree$tip.label<-gsub("heloisafmnh", "heloisa", brilliant_tree$tip.label) #Atthis_heloisafmnh
brilliant_tree$tip.label<-gsub("longirost", "longirostris", brilliant_tree$tip.label) #Heliomaster_longirost
brilliant_tree$tip.label<-gsub("polytmuscjc", "polytmus", brilliant_tree$tip.label) #Trochilus_polytmuscjc
brilliant_tree$tip.label<-gsub("macroHM", "macrocerca", brilliant_tree$tip.label) #Hylonympha_macroHM
brilliant_tree$tip.label<-gsub("lugens_", "lugens", brilliant_tree$tip.label)#Haplophaedia_lugens_
brilliant_tree$tip.label<-gsub("pella_", "pella", brilliant_tree$tip.label)#Topaza_pella_
brilliant_tree$tip.label<-gsub("cyanoJM", "cyanopectus", brilliant_tree$tip.label) #Sternoclyta_cyanoJM
brilliant_tree$tip.label<-gsub("Caprimul_longirostrisrisB", "Systellura_longirostris", brilliant_tree$tip.label)#Caprimul_longirostrisB
brilliant_tree$tip.label<-gsub("Streptop_zonaris_NMNHb", "Streptoprocne_zonaris", brilliant_tree$tip.label)#Streptop_zonaris_NMNHb
brilliant_tree$tip.label<-gsub("Campylopt_hemileucurus", "Campylopterus_hemileucurus", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Sternoclyta_cyanoIC", "Sternoclyta_cyanopectus", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Glaucis_hirsuta", "Glaucis_hirsutus", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Anthrac_nigricollis", "Anthracothorax_nigricollis", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Lodiggesia_mirabilis", "Loddigesia_mirabilis", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Eriocnemis_vestita", "Eriocnemis_vestita", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Boissonn_flavescens", "Boissonneaua_flavescens", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Urosticte_benjamani", "Urosticte_benjamini", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Coeligena_lutetia", "Coeligena_lutetiae", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Urochroa_bouguerizmc", "Urochroa_bougueri", brilliant_tree$tip.label)
brilliant_tree$tip.label<-gsub("Oreotroch estella", "Oreotrochilus_estella", brilliant_tree$tip.label)

brilliant_tree$tip.label

brilliant_tree <- keep.tip(brilliant_tree, tbnames)
#brilliant_tree$tip.label  <- sub(" ", "_", brilliant_tree$tip.label)
brilliant_tree$tip.label
#looks good but I'm sure additional changes will be necessary to match with shapefile names
```


### Trim massive dataset to keep only birds in Brilliant phylogeny and data  variables we will use.
```{r}
#filter by breeding season If you want
# botw.res <-data[data$SEASONAL==1,] # Resident ranges
# botw.bre <- data[data$SEASONAL==2,] # Breeding ranges
# 
# brill.res <- botw.res%>%
#   filter(SCINAME %in% brilliant_tree$tip.label)
# 
# brill.bre <- botw.bre%>%
#   filter(SCINAME %in% brilliant_tree$tip.label)

brill.all <-data2%>%
  filter(SCINAME %in% tbnames)

# Which species are not in shape files?
# species <- as.data.frame(brilliant_tree$tip.label)
# not.in <- species%>%
#   filter(!`brilliant_tree$tip.label` %in% data2$SCINAME)
# 
# data2%>%
#   filter(str_detect(SCINAME, "Oreotroch"))


brill.all<-sf:::as_Spatial(brill.all)

# brill.bre<- as_Spatial(brill.bre)
# 
# brill.res<- as_Spatial(brill.res)

crs(brill.all) <-NA
# crs(brill.res) <-NA
# crs(brill.bre) <- NA
#write these to shapefiles

colnames(brill.all@data) <- c("binomial", "presence", "origin", "seasonal")
head(brill.all)

brill.all@data$binomial <- as.factor(brill.all@data$binomial)

# writeSpatialShape(brill.res, fn="mapping/brill_res.shp")
# writeSpatialShape(brill.bre, fn="mapping/brill_bre.shp")
writeSpatialShape(brill.all, fn="mapping/brill_all.shp")

brill.all <- readShapePoly(fn="mapping/brill_res")

#Check if same class as example data
class(brill.all)
class(Phyllomedusa)

```

# Create matrix of presence absence using`letsR` package
```{r}

#This does not work for me. BUT I CAN RUN WITHIN THE FUNCTION AND EXTRACT PERTINENT RESULTS
brill.mat <- lets.presab(brill.all, xmn = -88, xmx =-30, ymn=-58, ymx=17, resol = 0.1)
#NEED TO OPEN FUNCTION AND RUN PIECE BY PIECE#This is result

save(brill.mat, file="mapping/brill_pa.R")

#Extract matrix and save separately as well
Brill_PA_Mat <- brill.mat$Presence_and_Absence_Matrix
write.table(Brill_PA_Mat, file="data/Brill_PA_Mat_.txt", sep="\t")

plot(brill.mat)
```